---
title: "igj"
author: "Peter"
date: "`r Sys.Date()`"
output: html_document
---



```{r}
amir_deals <- readRDS("/Users/peterdunson/Downloads/seller_1.rds")
world_happiness_sugar <- readRDS("/Users/peterdunson/Downloads/world_happiness_sugar.rds")
```


```{r}
food_consumption
```


```{r}
belgium_consumption <- food_consumption %>%
  filter(country == "Belgium")
```

```{r}
usa_consumption <- food_consumption %>%
  filter(country == "USA")
```

```{r}
mean(belgium_consumption$consumption)
median(belgium_consumption$consumption)
```

```{r}
mean(usa_consumption$consumption)
median(usa_consumption$consumption)
```

```{r}
food_consumption %>%
  filter(country %in% c("Belgium", "USA")) %>%
  group_by(country) %>%
  summarize(mean_consumption = mean(consumption), median_consumption = median(consumption))
```

```{r}
food_consumption %>%
  filter(food_category == "rice")
```


```{r}
food_consumption %>%
  filter(food_category == "rice")
```

```{r}
ggplot(food_consumption, aes(x = co2_emission)) +
  geom_histogram() #didnt work how it was supposed to, added the df, aes(x = var) and worked
```

```{r}
food_consumption %>%
  filter(food_category == "rice") %>%
  summarize(mean_co2 = mean(co2_emission), median_co2 = median(co2_emission))
```

```{r}
dists <- msleep$sleep_total - mean(msleep$sleep_total)
dists
```

```{r}
squared_dists <- (dists)^2
```

```{r}
sum_sq_dists <- sum(squared_dists)
sum_sq_dists
```

```{r}
sum_sq_dists/82 #n-1
```

```{r}
var(msleep$sleep_total) #one funciton, variability
```

```{r}
sqrt(var(msleep$sleep_total)) #standard deviation
sd(msleep$sleep_total) #or LITERALLY JUST THIS, DUMB BITCH
```

```{r}
mean(abs(dists)) #mean absolute deviation, SD penalizes longer distances more than shorter, MAD penalizes distances equally
```

```{r}
quantile(msleep$sleep_total) #quartiles
```

```{r}
ggplot(msleep, aes(y = sleep_total)) +
  geom_boxplot()
```

```{r}
quantile(msleep$sleep_total, probs = c(0, 0.2, 0.4, 0.6, 0.8, 1)) #quantiles
```

```{r}
quantile(msleep$sleep_total, probs = seq(0, 1, 0.2)) #same as above
```

```{r}
quantile(msleep$sleep_total, .75) - quantile(msleep$sleep_total, .25) #IQR
```

```{r}
iqr <- quantile(msleep$bodywt, .75) - quantile(msleep$bodywt, .25)
lower_threshold <- quantile(msleep$bodywt, .25) - 1.5 * iqr
upper_threshold <- quantile(msleep$bodywt, .75) + 1.5 * iqr
```

```{r}
msleep %>% filter(bodywt < lower_threshold | bodywt > upper_threshold) %>% #note the odd use of "|" between things, cant be replaced with a comma...
  select(name, vore, sleep_total, bodywt)
```


```{r}
msleep
```

```{r}
quantile(food_consumption$co2_emission) #df first, then column name
```

```{r}
quantile(food_consumption$co2_emission, probs = seq(0, 1, .2))
```

```{r}
quantile(food_consumption$co2_emission, probs = seq(0, 1, .1))
```

```{r}
food_consumption %>%
  group_by(food_category) %>%
  summarize(var_co2 = var(co2_emission), sd_co2 = sd(co2_emission))
```

```{r}
ggplot(food_consumption, aes(co2_emission)) +
  geom_histogram() +
  facet_wrap(~ food_category) #not sure what this is still, some way to make multiple graphs based on type of food
```

```{r}
food_consumption
```

```{r}
emissions_by_country <- food_consumption %>%
  group_by(country) %>%
  summarize(total_emission = sum(co2_emission))

emissions_by_country
```

```{r}
q1 <- quantile(emissions_by_country$total_emission, .25)
q3 <- quantile(emissions_by_country$total_emission, .75)
iqr <- q3 - q1
q1
```

```{r}
lower <- q1 - 1.5 * iqr
upper <- q3 + 1.5 * iqr
```

```{r}
emissions_by_country %>%
  filter(total_emission < lower | total_emission > upper)
```

```{r}
amir_deals
```

```{r}
amir_deals %>%
  count(product) %>%
  mutate(prob = n/sum(n))
```

```{r}
set.seed(31)

amir_deals %>%
  sample_n(5)
```

```{r}
set.seed(31)

amir_deals %>%
  sample_n(5, replace = TRUE)
```

```{r}
#imagine this is working, they wont let me fucking download the df, this is on DISCRETE DISTRIBUTIONS, picture of what DF looks like on ipad
ggplot(restaurant_groups, aes(group_size)) +
  geom_histogram(bins = 5)

size_distribution <- restaurant_groups %>%
  count(group_size) %>%
  mutate(probability = n / sum(n))

expected_val <- sum(size_distribution$group_size * size_distribution$probability)
```

```{r}
#prob of picking group of 4 or more
size_distribution %>%
  filter(group_size >= 4) %>%
  summarize(prob_4_or_more = sum(probability))
```



```{r}
punif(7, min = 0, max = 12) #prob of waiting less than or equal to 7
```

```{r}
punif(7, min = 0, max = 12, lower.tail = FALSE) #prob of more than or equal to 7
```

```{r}
punif(7, min = 0, max = 12) - punif(4, min = 0, max = 12) #prob 4-7
```

```{r}
min <- 0
max <- 30
```

```{r}
prob_less_than_5 <- punif(5, min = 0, max = 12)
```

```{r}
prob_greater_than_5 <- punif(5, min, max, lower.tail = FALSE)
```

```{r}
prob_between_10_20 <- punif(20, min, max) - punif(10, min, max)
```

```{r}
set.seed(334)
```

```{r}
wait_times %>%
  mutate(time = runif(1000, min = 0, max = 30)) #generating 1000 wait times between 0 and 30

ggplot(aes(time)) +
  geom_histogram(bins = 30)
```

```{r}
# rbinom(# of trials, # of coins, # probability of heads/success)
```


```{r}
coin_outcomes <- rbinom(100000, 1, 0.5)
```

```{r}
coin_outcomes <- rbinom(100000, 1, 0.5) #me messing around, not related
hist(coin_outcomes) #does this use ggplot?
```

```{r}
#me messing around, not related
ggplot(data = data.frame(outcomes = coin_outcomes), aes(x = outcomes)) +
  geom_histogram()

#The data argument specifies the data frame where the outcomes variable is stored, and the aes function defines the aesthetic mappings. In this case, x = outcomes specifies that the outcomes variable should be plotted on the x-axis.
#shocking that it knows what to do with 
```

```{r}
rbinom(8, 1, 0.5)
```

```{r}
rbinom(10, 3, 0.5) #represents the total number of heads from each flip, 10 flips, 3 coins, each coin is flipped 3 times...
```

```{r}
rbinom(10, 3, .25)
```

```{r}
#dbinom(num heads, num trials, prob of heads)
dbinom(7, 10, .5) #this is looking for the probability of getting heads 7times out of 10
```

```{r}
pbinom(7, 10, 0.5) #probability of 7 or fewer heads
```

```{r}
pbinom(7, 10, 0.5, lower.tail = FALSE) #p(heads > 7)
```

```{r}
set.seed(10)

rbinom(1, 1, .3) #single deal simulation
```

```{r}
rbinom(1, 3, .3)
```

```{r}
set.seed(10)

deals <- rbinom(52, 3, .3)

mean(deals)
```

```{r}
dbinom(3, 3, .3) #prob of 3/3
```

```{r}
pbinom(1, 3, .3) #prob less than one deal
```

```{r}
pbinom(1, 3, .3, lower.tail = FALSE) #prob of more than one deal
```

```{r}
won_30pct <- 3 * .3
won_25pct <- 3 * .25
won_35pct <- 3 * .35
```

NORMAL DISTR

```{r}
pnorm(154, mean = 161, sd = 7) #area of normal distr less than number, 154 in this case, percent of x lower than 154
```

```{r}
pnorm(154, mean = 161, sd = 7, lower.tail = FALSE) #inverse of above
```

```{r}
pnorm(157, mean = 161, sd = 7) - pnorm(154, mean = 161, sd = 7) #prob of between 154-157
```

```{r}
qnorm(.9, mean = 161, sd = 7) #what x is 90% less than, 90% is less than 169
```

```{r}
qnorm(.9, mean = 161, sd = 7, lower.tail = FALSE) #inverse of above
```

```{r}
rnorm(10, mean = 161, sd = 7) # 10 random from the normal distr
```

```{r}
ggplot(amir_deals, aes(amount)) +
  geom_histogram(bins = 10)
```

```{r}
pnorm(7500, mean = 5000, sd = 2000)
```

```{r}
pnorm(1000, mean = 5000, sd = 2000, lower.tail = FALSE)
```

```{r}
pnorm(7000, mean = 5000, sd = 2000) - pnorm(3000, mean = 5000, sd = 2000)
```

```{r}
qnorm(.75, mean = 5000, sd = 2000, lower.tail = FALSE) #75% of sales will be more than
```

```{r}
new_mean <- 1.2 * 5000
new_sd <- 1.3 * 2000
```

```{r}
new_sales <- new_sales %>% #this is what they gave and its just WRONG
  mutate(amount = rnorm(36, mean = new_mean, sd = new_sd))
```

```{r}
ggplot(new_sales, aes(amount)) + #shit dont work
  geom_histogram(bins = 10)
```

```{r}
pnorm(1000, mean = 5000, sd = 2000)
```

```{r}
pnorm(1000, mean = 6000, sd = 2600)
```

```{r}
die <- c(1, 2, 3, 4, 5, 6)
sample_of_5 <- sample(die, 5, replace = TRUE) #samples from vector rather than df (sample_n)
sample_of_5
```

```{r}
mean(sample_of_5)
```

```{r}
sample(die, 5, replace = TRUE) %>%
  mean()
```

```{r}
sample_means <- replicate(1000, sample(die, 5, replace = TRUE) %>%
  mean()) #10 diff sample means, be sure to include pipe and mean in parenthesis

sample_means #proof of CLT
```

```{r}
ggplot(data = data.frame(sample_means), aes(x = sample_means)) +
  geom_histogram()
```

```{r}
sample_sd <- replicate(1000, sample(die, 5, replace = TRUE) %>%
  sd()) #sd instead of means
```

```{r}
ggplot(data = data.frame(sample_sd), aes(x = sample_sd)) +
  geom_histogram()
```

```{r}
mean(sample_means)
```

```{r}
mean(sample_sd)
```

```{r}
ggplot(amir_deals, aes(num_users)) +
  geom_histogram(bins = 10)
```

```{r}
set.seed(104)

sample(amir_deals$num_users, 20, replace = TRUE) %>% #pay special attention to ad$nu... $$$$$$$$
  mean()
```

```{r}
sample_means <- replicate(100, sample(amir_deals$num_users, 20, replace = TRUE) %>%
  mean())
```

```{r}
ggplot(data = data.frame(sample_means), aes(x = sample_means)) +
  geom_histogram(bins = 10) #figured this out myself, datacamp some bums
```

```{r}
set.seed(321)

sample_means <- replicate(30, sample(amir_deals$num_users, 20) %>%
                            mean())
mean(sample_means)

mean(amir_deals$num_users)
```

poisson, lambda is expected value, distr peak is at lambda

```{r}
dpois(5, lambda = 8) #average is 8, what is P(5)
```

```{r}
ppois(5, lambda = 8) #P(<= 5)
```

```{r}
ppois(5, lambda = 8, lower.tail = FALSE)
```

```{r}
rpois(10, lambda = 8) #sample from poisson distr, CLT still applies
```

```{r}
dpois(5, lambda = 4)
```

```{r}
ppois(2, lambda = 4)
```

```{r}
ppois(10, lambda = 4, lower.tail = FALSE)
```

exponential distr, lambda = rate, continuous

```{r}
pexp(1, rate = 0.5) #P(less than a minute given rate is .5/min)
```

```{r}
pexp(4, rate = .5, lower.tail = FALSE) #P(waiting more than 4 min, given rate is .5/min)
```

```{r}
pexp(4, rate = .5) - pexp(1, rate = .5) #P(btwn 1 and 4)
```

1 / lambda = expected value for exp distr

t distr, thicker tails than normal, more likely further from mean, degrees of freedom

log normal distr, skewed distr

practice:
```{r}
pexp(1, rate = 1/2.5) #prob amir takes less than an hour to respond to lead, av=2.5
```

```{r}
pexp(4, rate = 1/2.5, lower.tail = FALSE)
```

```{r}
pexp(4, rate = 1/2.5) - pexp(3, rate = 1/2.5) #between 3 and 4 hrs
```

CORRELATION

Pearson product-moment correlation

```{r}
ggplot(df, aes(x, y)) +   #base model for visualizing relationships via scatter plots
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) #added trend line, lm = linear trend line, se false means no error margins
```

```{r}
cor(df$x, df$y)       #computing correlations, order doesn't matter, missing values means NA
```

```{r}
cor(df$x, df$y, use = "pairwise.complete.obs")     #to ignore missing values
```



```{r}
ggplot(world_happiness_sugar, aes(life_exp, happiness_score)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
cor(world_happiness_sugar$life_exp, world_happiness_sugar$happiness_score)
```

```{r}
cor(msleep$bodywt, msleep$awake)
```

LOG TRANSFORMATIONS, for skewed data

```{r}
msleep %>%
  mutate(log_bodywt = log(bodywt)) %>%
  ggplot(aes(log_bodywt, awake)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
cor(msleep$log_bodywt, msleep$awake) #didnt work...
```


```{r}
msleep_modified <- msleep %>%
  mutate(log_bodywt = log(bodywt)) %>% #CHAT GPT ADVICE, normal thing didnt work
  {cor(.$log_bodywt, .$awake)}
```

```{r}
msleep_modified
```

log(x)
sqrt(x)
1/x (reciprocal)
combinations of these things for x and y

correlation doesnt imply causation


CONFOUNDING VARS, association vs causation
```{r}
ggplot(world_happiness_sugar, aes(gdp_per_cap, life_exp)) +
  geom_point() 
```

```{r}
cor(world_happiness_sugar$gdp_per_cap, world_happiness_sugar$life_exp)
```

```{r}
ggplot(world_happiness_sugar, aes(happiness_score, gdp_per_cap)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)
```

```{r}
cor(world_happiness_sugar$happiness_score, world_happiness_sugar$gdp_per_cap)
```



```{r}
world_happiness <- world_happiness_sugar %>%
  mutate(log_gdp_per_cap = log(gdp_per_cap))
```

```{r}
ggplot(world_happiness, aes(happiness_score, log_gdp_per_cap)) +
  geom_point()
```

```{r}
cor(world_happiness$log_gdp_per_cap, world_happiness$happiness_score)
```

```{r}
ggplot(world_happiness_sugar, aes(grams_sugar_per_day, happiness_score)) +
  geom_point()
```

```{r}
cor(world_happiness_sugar$grams_sugar_per_day, world_happiness_sugar$happiness_score)
```

Designs of experiments

```{r}

```











